How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-28 tse-paper/content.tex
some fixes I merged away

% TODO Moritz: is this okay (here)?
% maybe it's better to add it to Threats to Validity with the heading:
% \texbf{Modifying the Build Process.}
An alternative to analyzing build logs, would be to integrate plugins
that report precise data about the build process, such as the JUnit
plugin for Jenkins~\cite{jenkins2020junit-plugin}.
However, such plugins are not available for all build tools and they
do not enable the analysis of historical data---builds that ran before
the plugin was installed cannot be analyzed.
In addition to that, when using such a plugin, one would have to know
beforehand which information should be reported from a build, whereas
build log analysis enables us to analyze the wide range of information
already baked into the build logs.
Their advantage is that they can precisely report the information which
is required.
A downside of plugins, which are integrated into the CI flow similar to
Deflaker by Bell et al.~\cite{bell2018deflaker}, is that they are unable
to work on historic
builds, i.e., before the moment they were being deployed and that if
information has not been recorded from the beginning but turns out to
be important, it is impossible to retrieve later.

@@

% Alt:
%\section{Analyzing Build Logs with Chunk Retrieval Techniques}
\section{Chunk Retrieval Techniques}
\label{sec:techniques}
%TODO: really? currently the discussion does not show this 'need'!
The outcome of our literature survey shows the need for a new
generation of techniques to analyze build logs.
In this section we argue how our research sets the first step to
address this need by introducing three techniques that retrieve
specific substrings (chunks) from build logs.
We present defining characteristics of these so called \emph{chunk
retrieval techniques} and present the techniques we developed in
detail.
% TODO At the end of what?
At the end we illustrate how the techniques operate on a concrete
example and sketch out their implementation.

@@

To fully comprehend how difficult the task of build log analysis really
is without any a-priori knowledge, we include a comparison with a
base line of
picking lines at random from the build log (RLR).
RLR mimics the
situation of guessing blindly which lines are interesting.
The only configuration option for RLR is the number of lines it should
retrieve.
For a fair comparison to the other techniques (whose number of
returned lines is dynamic), we configure RLR to return the average
number of lines in the chunks of the training examples.
giving it a slight advantage over guessing without previous knowledge.

@@

% TODO: want to?
% real-world log LogChunks ?
To illustrate the chunk retrieval techniques further, we want to give
an example of how the chunk retrieval techniques retrieve chunks from
real-world build logs.
\Cref{lst:chunk-example} shows two excerpts
from Android build logs.
Marked in blue are the chunks
which contain the information on why the
corresponding build failed.

@@

\begin{subfigure}[b]{\columnwidth}
  \begin{lstlisting}[breaklines=true,frame=tlr]
[0K$ test/sass-compile-tester.sh
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[backgroundcolor=\color{Cerulean!60},breaklines=true,frame=rl]
Error: Invalid CSS after "2.3em": expected expression (e.g.
1px, bold), was ";"
	on line 86 of sass/components/dropdown.sass
	from line 5 of sass/components/_all.sass
	from line 6 of bulma.sass
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[breaklines=true,frame=blr]
  Use --trace for backtrace.
  \end{lstlisting}
	\caption{Blue: targeted chunk ($DesiredLines$)}
	\label{lst:pbe-part-success-desired}
\end{subfigure}
  \caption{Example of a partially successful retrieval
  (PBE retrieved only
  two of the four targeted lines).}
  % TODO: shouldn't the label be partially successful? ;-)
  \label{lst:pbe-unsuccessful}
\end{figure}

@@

% actually, we could cut this figure and move it's explanation
% to the next one (this is an aggregation of the next one)
% however, having them separate eases the entry into the results a bit
% and we have explanation of the categories before the more complicated
% aspect of the change with more structural categories
\Cref{fig:failure-reason-PBE} shows the results of chunk
retrieval with PBE.
It shows how many of the retrieval runs were successful or
partially successful, as well as the number of runs where the
synthesized program did not match on the analyzed build log or where
no program could be synthesized.
% The last two results stem from the regular expressions and their
% synthesis, therefore only apply to PBE.
Out of the overall 400 runs, 5 per each one of the 80 example
sets, PBE extracted all desired lines successfully in 138 cases.
In 18 cases, the synthesized program only extracted a
subset of the desired lines, while in 12 cases it extracted none of
the desired lines.
For these 18 partially successful, the average recall
was 47\%.
\Cref{lst:pbe-unsuccessful} shows how an example of such
a partially successful retrieval looks like.
In it, the synthesized program only
retrieved two (\Cref{lst:pbe-part-success-output})
of the four (\Cref{lst:pbe-part-success-desired}) targeted lines.
In 59 cases, PBE could synthesize a regular expression program, though
it did not match on the test build log.
In 173 of the 400 cases
could PROSE not synthesize a program consistent with all of the training
examples.

>>>>

% TODO Moritz: is this okay (here)?
% maybe it's better to add it to Threats to Validity with the heading:
% \texbf{Modifying the Build Process.}
An alternative to analyzing build logs, would be to integrate plugins
that report precise data about the build process, such as the JUnit
plugin for Jenkins~\cite{jenkins2020junit-plugin}.
However, such plugins are not available for all build tools and they
do not enable the analysis of historical data---builds that ran before
the plugin was installed cannot be analyzed.
In addition to that, when using such a plugin, one would have to know
beforehand which information should be reported from a build, whereas
build log analysis enables us to analyze the wide range of information
already baked into the build logs.
Their advantage is that they can precisely report the information which
is required.

% TODO merge nicely
A downside of plugins, which are integrated into the CI flow similar to
Deflaker by Bell et al.~\cite{bell2018deflaker}, is that they are unable
to work on historic
builds, i.e., before the moment they were being deployed and that if
information has not been recorded from the beginning but turns out to
be important, it is impossible to retrieve later.

@@

% Alt:
%\section{Analyzing Build Logs with Chunk Retrieval Techniques}
\section{Chunk Retrieval Techniques}
\label{sec:techniques}
%TODO: really? currently the discussion does not show this 'need'!
The outcome of our literature survey shows the need for a new
generation of techniques to analyze build logs.
In this section we argue how our research sets the first step to
address this need by introducing three techniques that retrieve
specific substrings (chunks) from build logs.
We present defining characteristics of these so called \emph{chunk
retrieval techniques} and present the techniques we developed in
detail.
At the end of this section we illustrate how the techniques
operate on a concrete
example and sketch out their implementation.

@@

To fully comprehend how difficult the task of build log analysis really
is without any a-priori knowledge, we include a comparison with a
base line of
picking lines at random from the build log.
RLR mimics the
situation of guessing blindly which lines are interesting.
The only configuration option for RLR is the number of lines it should
retrieve.
For a fair comparison to the other techniques (whose number of
returned lines is dynamic), we configure RLR to return the average
number of lines in the chunks of the training examples.
giving it a slight advantage over guessing without previous knowledge.

@@

To illustrate the chunk retrieval techniques further, we give
an example of how the chunk retrieval techniques retrieve chunks from
real-world build logs.
\Cref{lst:chunk-example} shows two excerpts
from Android build logs.
Marked in blue are the chunks
which contain the information on why the
corresponding build failed.

@@

\begin{subfigure}[b]{\columnwidth}
  \begin{lstlisting}[breaklines=true,frame=tlr]
[0K$ test/sass-compile-tester.sh
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[backgroundcolor=\color{Cerulean!60},breaklines=true,frame=rl]
Error: Invalid CSS after "2.3em": expected expression (e.g.
1px, bold), was ";"
	on line 86 of sass/components/dropdown.sass
	from line 5 of sass/components/_all.sass
	from line 6 of bulma.sass
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[breaklines=true,frame=blr]
  Use --trace for backtrace.
  \end{lstlisting}
	\caption{Blue: targeted chunk ($DesiredLines$)}
	\label{lst:pbe-part-success-desired}
\end{subfigure}
  \caption{Example of a partially successful retrieval
  (PBE retrieved only
  two of the four targeted lines).}
  \label{lst:pbe-partially-successful}
\end{figure}

@@

% actually, we could cut this figure and move it's explanation
% to the next one (this is an aggregation of the next one)
% however, having them separate eases the entry into the results a bit
% and we have explanation of the categories before the more complicated
% aspect of the change with more structural categories
\Cref{fig:failure-reason-PBE} shows the results of chunk
retrieval with PBE.
It shows how many of the retrieval runs were successful or
partially successful, as well as the number of runs where the
synthesized program did not match on the analyzed build log or where
no program could be synthesized.
% The last two results stem from the regular expressions and their
% synthesis, therefore only apply to PBE.
Out of the overall 400 runs, 5 per each one of the 80 example
sets, PBE extracted all desired lines successfully in 138 cases.
In 18 cases, the synthesized program only extracted a
subset of the desired lines, while in 12 cases it extracted none of
the desired lines.
For these 18 partially successful, the average recall
was 47\%.
\Cref{lst:pbe-partially-successful} shows how an example of such
a partially successful retrieval looks like.
In it, the synthesized program only
retrieved two (\Cref{lst:pbe-part-success-output})
of the four (\Cref{lst:pbe-part-success-desired}) targeted lines.
In 59 cases, PBE could synthesize a regular expression program, though
it did not match on the test build log.
In 173 of the 400 cases
could PROSE not synthesize a program consistent with all of the training
examples.
